---
title: "Evolution stochasticity converts additive tradeoffs to sub-additive"
author: "Lucas A. Nell"
date: "9/14/2020"
tables: true
graphics: true
colorlinks: false
mathspec: true
linestretch: 2
output: 
  bookdown::pdf_document2: 
    highlight: haddock
    toc: false
    numbered_sections: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
    library(knitr)
    library(tidyverse)
    library(sauron)
    library(cowplot)
    library(egg)
})

knitr::opts_chunk$set(echo = FALSE, cache = FALSE, eval = TRUE, 
                      warning = FALSE, message = FALSE, 
                      dev = "cairo_pdf")
theme_set(theme_classic() + 
              theme(legend.background = element_blank(),
                    strip.background = element_blank(),
                    plot.margin = margin(0.5, 0.5, 0.5, 2),
                    strip.text = element_text(size = 10),
                    panel.spacing = unit(0.5, "lines")))

if (!isTRUE(getOption('knitr.in.progress'))) {
    options("device" = "quartz")
    grDevices::graphics.off()
}
options(dplyr.summarise.inform = FALSE)
# Clean captions
cc <- function(.x) {
    .x <- gsub("\n", " ", .x)
    .x <- gsub("\\s+", " ", .x)
    return(.x)
}
```



## Introduction

Here, I describe the issue I'm having when tradeoffs are additive and 
we add stochasticity to evolution.
Without stochasticity, species evolve to have any axis-value
combination that yields a position on the neutral ring.
With it, species evolve to have intermediate
values of both axes (i.e., they evolve to the center of the neutral ring).
This is similar to what happens when tradeoffs are sub-additive.
In fact, the axis values that species evolve to under this situation appear to 
be the values that are approached as $\eta$ approaches zero from the negative
side.

My goal here is to outline the weirdness of this issue and ask what you
recommend I do about it in the paper.
I'm just not sure how to talk about this in the paper, as it appears to
be an artifact of plugging the random variables into nonlinear equations.


## Equations


This issue happens regardless of the number of species, so I'm
using only one so that we can ignore conflicting vs ameliorative evolution
(i.e., the $\mathbf{D}$ matrix).
First, here are the relevant equations for the one-species, two-axis case:

\begin{equation*}
\begin{split}
    F_{t+1} &= \exp \left\{
        r_0 - f ( \ddot{v}_{1,t}^2 + 2 \eta \, \ddot{v}_{1,t} \, \ddot{v}_{2,t} + \ddot{v}_{2,t}^2 ) -
        \alpha_0 \, \textrm{e}^{ - \ddot{v}_{1,t}^2 - \ddot{v}_{2,t}^2 } \, N_t
    \right\} \\
    \ddot{v}_{k,t+1} &= v_{k,t+1} \; \text{e}^{\varepsilon_V} \\
    \varepsilon_V &\sim \text{N}( 0, \, \sigma^2_V) \\
    v_{1,t+1} &= v_{1,t} + 2 \; \sigma^2
    \left[
        \alpha_0 \; N_t \;
            \textrm{e}^{-\ddot{v}_{1,t}^2 - \ddot{v}_{2,t}^2} \; \ddot{v}_{1,t}
        - f \; ( \ddot{v}_{1,t} + \eta \; \ddot{v}_{2,t} )
    \right] \\
    v_{2,t+1} &= v_{2,t} + 2 \; \sigma^2
    \left[
        \alpha_0 \; N_t \;
            \textrm{e}^{-\ddot{v}_{2,t}^2 - \ddot{v}_{1,t}^2} \; \ddot{v}_{2,t}
        - f \; ( \ddot{v}_{2,t} + \eta \; \ddot{v}_{1,t} )
    \right] \\
    \textrm{.}
\end{split}
\end{equation*}


where we have the following parameters:


\begin{tabular}{ll}  
\toprule
Parameter    & Description \\
\midrule
$F$                 & fitness \\
$r_0$               & baseline growth rate \\
$f$                 & cost of increasing axes on the growth rate \\
$\ddot{v}_{k,t}$    & axis $k$ phenotype at time $t$ \\
$v_{k,t}$           & axis $k$ genotype at time $t$ \\
$\eta$              & non-additive tradeoff \\
$\alpha_0$          & baseline density dependence \\
$N$                 & population abundance \\
$\varepsilon_V$     & axis-evolution stochasticity \\
$\sigma^2_V$        & variance for axis-evolution stochasticity \\
$\sigma^2$          & additive genetic variance \\
\bottomrule
\end{tabular}



\textbf{\emph{Note:}} Changing the $\mu$ parameter for the lognormal distribution to
$\frac{ - \sigma^2_V }{ 2 }$ so that 
$\text{E}(\text{e}^{\varepsilon_V}) = 1$
has no effect.



## Selection

The figure below shows the direction and relative magnitude of selection
under the simulation scenario used throughout.


```{r parameters}

# d1 = d2 = 0
D <- matrix(0, 2, 2)

# eta = 0
C <- diag(2)

# Defaults from `quant_gen`
f <- formals(quant_gen)$f
a0 <- formals(quant_gen)$a0
r0 <- formals(quant_gen)$r0

# Equilibrium abundance for deterministic sims:
N <- 10.91963

```


```{r ss-fig-caption}
ss_cap <- "
Direction of selection strength for a two-axis, one species community
with additive tradeoffs.
Arrow length is proportional to the strength of selection for the axis
values at the base of the arrow.
The gray ring indicates the neutrally stable equilibrium.
"
```
```{r selection-strength, fig.width=4, fig.height=4, fig.cap=cc(ss_cap)}

dir_df <- crossing(.v1 = seq(0, 4, 0.4),
                   .v2 = seq(0, 4, 0.4)) %>%
    mutate(across(.fns = round, digits = 3)) %>%
    pmap_dfr(function(.v1, .v2) {
        ss <- sauron:::sel_str_cpp(rbind(.v1, .v2), N, f, a0, C, r0, D)
        tibble(V1 = .v1, V2 = .v2, V1_delta = ss[1,1], V2_delta = ss[2,1])
    })

dir_df %>%
    ggplot(aes(V1, V2)) +
    stable_points(0, return_geom = TRUE, color = "gray70") +
    geom_segment(aes(xend = V1 + 0.1 * V1_delta, yend = V2 + 0.1 * V2_delta),
                 arrow = arrow(length = unit(0.05, "inches"))) +
    scale_fill_viridis_c() +
    scale_x_continuous("Axis 1", breaks = c(0, 2, 4)) +
    scale_y_continuous("Axis 2", breaks = c(0, 2, 4)) +
    coord_equal() +
    NULL
```



## Stochasticity of various forms


In these simulations, there is just one species with two axes and non-additive
tradeoffs.
The species genotype starts on the upper part of the neutral ring, and
its abundance is the equilibrium abundance as calculated based on 
the default parameters (e.g., $f$, $r_0$, $\alpha_0$).
Abundance does not change in these simulations, as I wanted to remove that
as a potential part of the problem.
(It turns out not to affect the outcomes.)


I simulated stochasticity in two forms. 
The first is described above, where
phenotypes are the product of the genotype and a lognormal random variable.
This form is called "$\times$ lnorm" in figures below.
The second form of stochasticity is called "+ rnorm" because it's where
phenotypes are genotypes plus a normal random variable 
(truncated to be $\ge 0$).


I also simulated for high and low levels of stochasticity.
For the lognormal form of stochasticity, I chose $\sigma_V$ of 0.1 and 0.2.
I simulated the normal form multiple times to choose values of $\sigma_V$
that produced approximately the same standard deviation between genotypes
and phenotypes as for the lognormal form with $\sigma_V$ of 0.1 and 0.2.
The values chosen were 0.13 and 0.29.


These simulations show the problem for the lognormal form of stochasticity:
it inextricably moves toward the middle point on the ring
(Figure \@ref(fig:stoch-forms-sims-fig)A,B).
This is most pronounced for a high $\sigma_V$.
It doesn't just occur for the lognormal distribution, though.
The normal distribution moves the species toward the middle point, though
it's slower.
When I increased $\sigma_V$ to 0.5 for the normal distribution, it looked
pretty similar to the lognormal distribution at 0.2.
Thus the normal distribution does the same thing, but the deviance between
genotype and phenotype has to be greater for it to do so.


We can also see that both forms of stochasticity caused the distance from 
the origin to be greater than the ideal value 
(Figure \@ref(fig:stoch-forms-sims-fig)C).
I believe this is because selection pushes much stronger from the origin
to the ring than from outside the ring to the ring 
(Figure \@ref(fig:selection-strength)).
Thus when phenotypes are inside the ring, selection causes
genotypes to overshoot the ring.





```{r sim-function}


#' Function to run simulations
#'
#' @param .method Which method for stochasticity. `1`: genotype * lnorm,
#'     `2`: genotype + norm.
#' @param .sigma_V Sigma parameter for simulations. 
#'     Note that this isn't necessarily the resulting standard deviation.
#' @param .stoch Length-2 vector indicating which axes receive stochasticity.
#' @param maxt Number of simulation time steps.
#' @param .start Indicates starting axis values.
#'
Vstoch_sim <- function(.method,
                       .sigma_V,
                       .stoch = c(TRUE, TRUE),
                       maxt = 2e3,
                       .start = "over") {
    
    V_test <- matrix(0, maxt+1, 4)
    
    if (is.character(.start)) {
        .start <- match.arg(.start, c("over", "under"))
        if (.start == "over") {
            V_test[1,1:2] <- c(0.4, sqrt(4-0.4^2))
        } else V_test[1,1:2] <- c(sqrt(4-0.4^2), 0.4)
    } else if (is.numeric(.start) && length(.start) == 2) {
        V_test[1,1:2] <- .start
    } else stop(".start must be char or length-2 numeric")
    
    # stoch_fun <- list(function(x) x * rlnorm(2, - .sigma_V^2 / 2, .sigma_V),
    stoch_fun <- list(function(x) x * rlnorm(2, 0, .sigma_V),
                      function(x) trnorm(2, x, sqrt(exp(.sigma_V^2) - 1)))
    stoch_fun <- stoch_fun[[.method]]
    
    V_test[1,3:4] <- stoch_fun(V_test[1,1:2])
    
    if (!.stoch[1]) V_test[1,3] <- V_test[1,1]
    if (!.stoch[2]) V_test[1,4] <- V_test[1,2]
    
    for (t in 1:maxt) {
    
        deltaV <- 0.05 * sauron:::sel_str_cpp(cbind(V_test[t,3:4]), N, f, a0, 
                                              C, r0, D)
        
        V_test[t+1,1:2] <- pmax(0, V_test[t,1:2] + deltaV[,ncol(deltaV)])
        V_test[t+1,3:4] <- stoch_fun(V_test[t+1,1:2])
    
        if (!.stoch[1]) V_test[t+1,3] <- V_test[t+1,1]
        if (!.stoch[2]) V_test[t+1,4] <- V_test[t+1,2]
    
    }
    
    colnames(V_test) <- c("V1", "V2", "Vp1", "Vp2")
    
    out <- as_tibble(V_test) %>%
        mutate(time = (1:n()) - 1L) %>% 
        select(time, everything())
    
    return(out)
}

# Test for SD between phenotype and genotype
stoch_test <- function(Vstoch_sim_out) {
    n <- nrow(Vstoch_sim_out)
    c(sqrt(sum((Vstoch_sim_out[,"Vp1"] - Vstoch_sim_out[,"V1"])^2) / n),
      sqrt(sum((Vstoch_sim_out[,"Vp2"] - Vstoch_sim_out[,"V2"])^2) / n))
}


```

```{r stoch-forms-sims-fig-caption}
ss_fig1_cap <- "
Representative simulations for the change in
(A) axis space,
(B) angle from origin, and
(C) distance from origin
through time.
This is for high and low levels of stochasticity (sub-panel columns) and 
for lognormal and normal error distributions (rows).
(A) Circles represent starting genotype locations,
triangles are ending locations,
the gray arc is the neutrally stable ring, and 
the dotted line is the 1-to-1 line.
(All panels) Black lines represent paths of genotype change, and 
blue lines indicate paths of phenotype change.
(B) The red line indicates the angle for the middle point on the ring.
(C) The red line indicates the ideal radius (i.e., the one that maximizes 
fitness).
"
```


```{r stoch-forms-sims-fig, fig.width=6.5, fig.height=4, fig.cap=cc(ss_fig1_cap)}

# stoch_test(Vstoch_sim(2, 0.29))
# method 1 @ 0.2:  0.2438648 0.3382268
# method 1 @ 0.1:  0.07617966 0.19386198


sims1 <- crossing(.method = 1:2,
                  .sigma_V = c(0.1, 0.2)) %>% 
    pmap_dfr(function(.method, .sigma_V) {
        .s <- case_when(.method == 1 ~ .sigma_V,
                        .method == 2 & .sigma_V == 0.1 ~ 0.13,
                        .method == 2 & .sigma_V == 0.2 ~ 0.29,
                        TRUE ~ NA_real_)
        stopifnot(!is.na(.s))
        Vstoch_sim(.method, .s) %>% 
            mutate(method = .method, sigma_V = .sigma_V)
    }) %>%
    mutate(method = factor(method, levels = 1:2, labels = c("'× lnorm'", "'+ norm'")),
           sigma_V = factor(sigma_V, levels = sort(unique(sigma_V)),
                            labels = sprintf("%s ~ sigma[v]", c("low", "high"))))

ss_sims_fig1a <- sims1 %>% 
    ggplot(aes(V1, V2)) +
    geom_path(aes(Vp1, Vp2), color = "dodgerblue", alpha = 0.5) +
    geom_path() +
    stable_points(0, return_geom = TRUE, color = "gray40") +
    geom_point(data = sims1 %>% filter(time == 0),
               shape = 19, size = 4, color = "firebrick") +
    geom_point(data = sims1 %>% filter(time == max(time)),
               shape = 17, size = 4, color = "firebrick") +
    geom_abline(slope = 1, intercept = 0, linetype = 2, color = "gray70") +
    xlab("Axis 1") +
    ylab("Axis 2") +
    facet_grid(method ~ sigma_V, labeller = label_parsed) +
    coord_equal(xlim = c(0,4), ylim = c(0,4)) +
    theme(strip.text.y = element_text(angle = 0))


ss_sims_fig1b <- sims1 %>% 
    mutate(angle = atan(Vp1 / Vp2) * 180 / pi) %>%
    ggplot(aes(time / 1000, angle)) +
    geom_line(color = "dodgerblue", alpha = 0.5) +
    geom_line(data = sims1 %>%
                  mutate(angle = atan(V1 / V2) * 180 / pi),
              color = "black") +
    geom_hline(yintercept = 45, color = "firebrick") +
    geom_hline(yintercept = 0, color = "gray70") +
    facet_grid(method ~ sigma_V, labeller = label_parsed) +
    ylab("Angle (º)") +
    scale_x_continuous("Time (× 1000)", breaks = 0:2) +
    theme(strip.text.y = element_text(angle = 0))


ss_sims_fig1c <- sims1 %>% 
    mutate(radius = sqrt(Vp1^2 + Vp2^2)) %>%
    ggplot(aes(time / 1000, radius)) +
    geom_line(color = "dodgerblue", alpha = 0.5) +
    geom_line(data = sims1 %>%
                  mutate(radius = sqrt(V1^2 + V2^2)),
              color = "black") +
    geom_hline(aes(yintercept = 2), color = "firebrick") +
    facet_grid(method ~ sigma_V, labeller = label_parsed) +
    ylab("Distance from origin") +
    scale_x_continuous("Time (× 1000)", breaks = 0:2) +
    theme(strip.text.y = element_text(angle = 0))



ggarrange(ss_sims_fig1a + theme(strip.text.y = element_blank()), 
          ss_sims_fig1b + theme(strip.text.y = element_blank()), 
          ss_sims_fig1c, 
          nrow = 1, labels = c("A", "B", "C"), 
          label.args = list(gp = grid::gpar(fontface = "bold")))

# sims1 %>% 
#     filter(time > 100) %>% 
#     group_by(method, sigma_V) %>% 
#     summarize(radius = mean(sqrt(V1^2 + V2^2)))



```



## Stochasticity for one or another axis

These are similar simulations to the previous section, except that I 
kept $\sigma_V = 0.2$ and instead changed which axis I added stochasticity to.
The main result here is that whichever axis had stochasticity, 
the species evolved to have that axis be zero
(Figure \@ref(fig:stoch-axes-sims-fig)).




```{r stoch-axes-sims-fig-caption}
ss_fig2_cap <- "
Representative simulations for the change in
(A) axis space,
(B) angle from origin, and
(C) distance from origin
through time.
This is for axis 1 or 2 having stochasticity (sub-panel columns) and 
for lognormal and normal error distributions (rows).
$\\sigma_V = 0.2$ for all panels.
(A) Circles represent starting genotype locations,
triangles are ending locations,
the gray arc is the neutrally stable ring, and 
the dotted line is the 1-to-1 line.
(All panels) Black lines represent paths of genotype change, and 
blue lines indicate paths of phenotype change.
(B) The red line indicates the angle for the middle point on the ring.
(C) The red line indicates the ideal radius (i.e., the one that maximizes 
fitness).
"
```


```{r stoch-axes-sims-fig, fig.width=6.5, fig.height=4.5, fig.cap=cc(ss_fig2_cap)}


sims2 <- crossing(.method = 1:2,
                  .sigma_V = 0.2,
                  .stoch = list(c(TRUE, FALSE), c(FALSE, TRUE)),
                  .start = c("over", "under")) %>% 
    pmap_dfr(function(.method, .sigma_V, .stoch, .start) {
        .s <- case_when(.method == 1 ~ .sigma_V,
                        .method == 2 & .sigma_V == 0.1 ~ 0.13,
                        .method == 2 & .sigma_V == 0.2 ~ 0.29,
                        TRUE ~ NA_real_)
        stopifnot(!is.na(.s))
        Vstoch_sim(.method, .s, .stoch, .start = .start) %>% 
            mutate(method = .method, stoch = which(.stoch), start = .start)
    }) %>%
    mutate(method = factor(method, levels = 1:2, labels = c("× lnorm", "+ norm")),
           stoch = factor(stoch, levels = 1:2,
                            labels = sprintf("stoch. in\naxis %i", 1:2)),
           start = factor(start, levels = c("over", "under")))


..start <- "over"

ss_sims_fig2a <- sims2 %>%
    filter(start == ..start) %>% 
    ggplot(aes(V1, V2)) +
    geom_path(aes(Vp1, Vp2), color = "dodgerblue", alpha = 0.5) +
    geom_path() +
    stable_points(0, return_geom = TRUE, color = "gray40") +
    geom_point(data = sims2 %>% filter(time == 0, start == ..start),
               shape = 19, size = 4, color = "firebrick") +
    geom_point(data = sims2 %>% filter(time == max(time), start == ..start),
               shape = 17, size = 4, color = "firebrick") +
    geom_abline(slope = 1, intercept = 0, linetype = 2, color = "gray70") +
    xlab("Axis 1") +
    ylab("Axis 2") +
    facet_grid(method ~ stoch) +
    coord_equal(xlim = c(0,4), ylim = c(0,4)) +
    theme(strip.text.y = element_text(angle = 0))


ss_sims_fig2b <- sims2 %>%
    filter(start == ..start) %>% 
    mutate(angle = atan(Vp1 / Vp2) * 180 / pi) %>%
    ggplot(aes(time / 1000, angle)) +
    geom_line(color = "dodgerblue", alpha = 0.5) +
    geom_line(data = sims2 %>%
                  filter(start == ..start) %>% 
                  mutate(angle = atan(V1 / V2) * 180 / pi),
              color = "black") +
    geom_hline(yintercept = 45, color = "firebrick") +
    geom_hline(yintercept = 0, color = "gray70") +
    facet_grid(method ~ stoch) +
    ylab("Angle (º)") +
    scale_x_continuous("Time (× 1000)", breaks = 0:2) +
    theme(strip.text.y = element_text(angle = 0))


ss_sims_fig2c <- sims2 %>%
    filter(start == ..start) %>% 
    mutate(radius = sqrt(Vp1^2 + Vp2^2)) %>%
    ggplot(aes(time / 1000, radius)) +
    geom_line(color = "dodgerblue", alpha = 0.5) +
    geom_line(data = sims2 %>%
              filter(start == ..start) %>% 
              mutate(radius = sqrt(V1^2 + V2^2)),
              color = "black") +
    geom_hline(aes(yintercept = 2), color = "firebrick") +
    facet_grid(method ~ stoch) +
    ylab("Distance from origin") +
    scale_x_continuous("Time (× 1000)", breaks = 0:2) +
    theme(strip.text.y = element_text(angle = 0))



ggarrange(ss_sims_fig2a + theme(strip.text.y = element_blank()), 
          ss_sims_fig2b + theme(strip.text.y = element_blank()), 
          ss_sims_fig2c, 
          nrow = 1, labels = c("A", "B", "C"), 
          label.args = list(gp = grid::gpar(fontface = "bold")))


```
