---
title: "Disruptive selection"
author: "Lucas A. Nell"
date: "10/21/2018"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Disruptive selection}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "svglite", 
  fig.width = 6,
  fig.height = 4,
  echo = FALSE
)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(tibble.print_min = 4L, tibble.print_max = 8L)
```
```{r source_Rprofile, echo = FALSE}
if (!isTRUE(getOption('knitr.in.progress'))) source(".Rprofile")
suppressPackageStartupMessages(library(tidyverse))
```



## Continuous trait

Population growth is as follows:

$$
\begin{align}
N_{t + 1} &= N_t F(\nu_t, N_t) \\
F(r, N_t) &= 1 + r (1 - \frac{N_t}{K}) \\
r(\nu_t) &= r_0 \left( 1 - e^{- \nu_t^2} \right) \\
& \text{so...} \\\
F(\nu_t, N_t) &= 1 + r_0 \left[ 1 - e^{- \nu_t^2} \right] \left[ 1 - \frac{N_t}{K} \right] \\
\end{align}
$$
where $F$ is mean population fitness,
$\nu$ is the mean trait value,
$N$ is population size,
$K$ is carrying capacity,
and
$r_0$ is population growth rate.




From Northfield & Ives (2013; doi: 10.1371/journal.pbio.1001685):

$$
\begin{align}
\nu_{t+1} &= \nu_t + \frac{V}{F} \frac{\partial F}{\partial \nu} \\
\end{align}
$$
where $V$ is the additive genetic variance.


$$
\begin{align}
& \frac{\partial F}{\partial \nu} \left( 1 + r_0 \left[ 1 - e^{- \nu^2} \right] 
    \left[ 1 - \frac{N}{K} \right] \right) \\
&= -2 r_0 ~ e^{-\nu^2} \nu \left[ 1 - \frac{N}{K} \right] \\
\end{align}
$$

So...

$$
\begin{align}
\nu_{t+1} &= \nu_t + \frac{V}{F} \left\{ -2 r_0 ~ e^{-\nu_t^2} \nu_t \left[ 1 - \frac{N}{K} \right] \right\} \\
\end{align}
$$

In the above case, you can see that the mean trait value, assuming $V > 0$, reaches
an equilibrium when the population reaches its carrying capacity $N=K$
and when $\nu_t = 0$.



